---
title: "Diamonds"
author: "James Hamilton"
date: "March 20, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
data(diamonds)
```

### Scatterplot price vs. x

Ignored 8 items with x = 0, focused on X between 3.7 and 7.5 and
price between 0 and 10000.
```{r}
ggplot(data = subset(diamonds, x > 0),
       aes(x = x, y = price)) +
  geom_point(alpha = 1/100) +
  coord_cartesian(xlim = c(3.7,7.5), ylim = c(0,10000))
```

### Correlation tests price vs. (x, y, z)
```{r}
cor.test(diamonds$price, diamonds$x)
cor.test(diamonds$price, diamonds$y)
cor.test(diamonds$price, diamonds$z)
```

### Plot price vs. depth
```{r}
ggplot(data = diamonds, aes(x = depth, y = price)) +
  geom_point(alpha = 1/25) +
  coord_cartesian(xlim = c(57.5,65)) +
  scale_y_log10()
```

### Correlation of price vs. depth
```{r}
cor.test(diamonds$price, diamonds$depth)
```

### Plot price vs. carat - eliminating the top 1% of values from both variables
```{r}
filter(diamonds, price < quantile(diamonds$price, probs = c(.99))) %>%
  filter(carat < quantile(diamonds$carat, probs = c(.99))) %>%
  ggplot(aes(x = carat, y = price)) +
    geom_point(alpha = 1/20) +
    scale_y_log10() 
```

### Plot price vs. volume
First calculate the volume and save it, then make the plot with not mods yet.
```{r}
diamonds$volume <- with(diamonds, x * y * z)

ggplot(data = diamonds, aes(x = volume, y = price)) +
  geom_point()
```

### Correlation of price and volume
```{r}
corrdf <- with(diamonds, diamonds[ volume > 0 & volume <= 800, ])
with(corrdf, cor.test(price, volume, data = corrdf))
```

### New plot with subset of data
```{r}
ggplot(data = corrdf, aes(x = volume, y = price)) +
  geom_point(alpha = 1/20) + 
  stat_smooth(method = "lm", formula = y ~ x)
```

### Create diamondsByClarity using dplyr
```{r}
diamondsByClarity <- diamonds %>%
  group_by(clarity) %>%
  summarise(
    mean_price = mean(price),
    median_price = median(price),
    min_price = min(price),
    max_price = max(price),
    n = n()
  )
```

### Use grid.arrange to display multiple plots in one window.
```{r}
library(gridExtra)

diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

p1 <- ggplot(data = diamonds_mp_by_clarity, aes(x = clarity, y = mean_price)) +
  geom_bar(stat = "identity")

p2 <- ggplot(data = diamonds_mp_by_color, aes(x = color, y = mean_price)) +
  geom_bar(stat = "identity")

grid.arrange(p1, p2)
```





